---
title: "2016 US Election Results"
author: "Bill Martersteck"
date: "11/15/2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
options(warn = -1) 
```

## Electoral Votes

- US Election held on 11/8/2016
- Each US State and the District of Columbia are given a number of electoral votes to award
to the presidential candidate as defined by state law.
- Each state gets an electoral vote for each US Senator and each House of Representative member (for a minimum of 3). More populous states, therefore, award more electoral votes.
- The District of Columbia gets to award 3 Electoral Votes. This right was granted
by the 23rd Ammendment to the Constitution.
- Donald Trump (R) won the 2016 Presidential Election by winning a majority of electoral votes.

## 2016 State Electoral Votes and Election Results

```{r, echo=FALSE, message=FALSE}
# Create data frame
evdf <- read.csv("election.csv", stringsAsFactors=FALSE)
evdf$Declared = as.factor(evdf$Declared)
evdf$Winner = ifelse(evdf$Clinton.EV>evdf$Trump.EV,"Clinton","Trump")

# Create hover text
evdf$hover <- with(evdf, paste0(State, '<br>',
            sprintf("Population: %.2f M<br>", Population/(10^6)),
            sprintf("Electoral Votes: %i<br>", State.EV),
            sprintf("Clinton %% of Votes: %.2f<br>", (Clinton*100)/(Clinton+Trump+Others)),
            sprintf("Trump %% of Votes: %.2f<br>", (Trump*100)/(Clinton+Trump+Others)),
            "Clinton EV: ", ifelse(evdf$Declared=="Decided",evdf$Clinton.EV,"Undecided"), "<br>",
            "Trump EV: ", ifelse(evdf$Declared=="Decided",evdf$Trump.EV,"Undecided")
            ))

# Make state borders white
borders <- list(color = toRGB("white"))
# Set up some mapping options
map_options <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = FALSE
)

plot_ly(z = evdf$State.EV, text = evdf$hover, locations = evdf$St, 
        type = 'choropleth', locationmode = 'USA-states', 
        color = evdf$State.EV, marker = list(line = borders)) %>%
        layout(geo = map_options)


```
